Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-09-10 paper.tex
unfinished pegasus summary

\begin{document}
\title{Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data}
\author{Valerie Hayot-Sasson}
\maketitle
\begin{abstract}
  Text of abstract.  
\end{abstract}
\tableofcontents
\chapter{Introduction}
intro here -mention hpc, data centre, cloud. 
\chapter{BigData and Open science in neuroimaging}
    \note{Likely will not be split in two separate sections}
    \section{large images}
        \begin{itemize}  
            \item BigBrain
            \item Micro CT 
            \item EM 
        \end{itemize}
    \section{large datasets}
    	\note{off the top of my head, but need to double check}
        \begin{itemize}
            \item HCP
            \item UK Biobank
            \item ADNI
            \item OpenFMRI
            \item BIDS/BIDS apps
        \end{itemize}
    An effective engine to process such data requires the following characteristics
        \begin{itemize}
            \item in-memory computing
            \item data locality
            \item lazy evaluation
        \end{itemize}
        \section{open-science /data sharing platforms for neuroimaging}
        		is open science just data sharing, computing. 
        		\subsection{Introduction}
		\subsection{CBRAIN}
		\subsection{OpenNEURO}
		\subsection{NeuroData}
\chapter{Workflow engines for neuroimaging data}
	\section{Definition of workflows/pipelines}

@@

    \section{Pegasus}
        \subsection{Pipeline description}
        \subsection{Pipeline execution}
        \subsection{Debugging and provenance}
    \section{Geneflow}
        \subsection{Pipeline description}
        \subsection{Pipeline execution}
        \subsection{Debugging and provenance}
    \section{Toil}
        \subsection{Pipeline description}
        \subsection{Pipeline execution}
        \subsection{Debugging and provenance}
    \section{Thunder}

>>>>

\begin{document}
\title{Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data}
\author{Valerie Hayot-Sasson}
\maketitle
\begin{abstract}
  Text of abstract.  
\end{abstract}
\tableofcontents
\chapter{Introduction}
intro here -mention hpc, data centre, cloud. 
\chapter{BigData and Open science in neuroimaging}
    \note{Likely will not be split in two separate sections}
    \section{large images}
        \begin{itemize}  
            \item BigBrain
            \item Micro CT 
            \item EM 
        \end{itemize}
    \section{large datasets}
    	\note{off the top of my head, but need to double check}
        \begin{itemize}
            \item HCP
            \item UK Biobank
            \item ADNI
            \item OpenFMRI
            \item BIDS/BIDS apps
        \end{itemize}
    An effective engine to process such data requires the following characteristics
        \begin{itemize}
            \item in-memory computing
            \item data locality
            \item lazy evaluation
        \end{itemize}
        \section{open-science /data sharing platforms for neuroimaging}
        		is open science just data sharing, computing. 
        		\subsection{Introduction}
		\subsection{CBRAIN}
		\subsection{OpenNEURO}
		\subsection{NeuroData}
        \subsection{Apache Airvata}
\chapter{Workflow engines for neuroimaging data}
	\section{Definition of workflows/pipelines}

@@

    \section{Pegasus}
        Pegasus is a general-purpose scientific workflow engine that specializes in mapping abstract descriptions of workflows to resources for efficient processing. 
        Similarly to other workflows, a Pegasus workflow can be represented by a DAG. This DAG is described by the user in DAX format (Directed Acyclic graph in XML).
        As the DAX abstract workflow is resource-independent, Pegasus can convert the abstract workflow into an execution workflow that is catered to the resources available.

        \subsection{Pipeline description}
        Within the DAX, the nodes are the tasks and the execution order is represented by the edges.
        In the instance of nested graphs, the nodes in the DAX may also represent a DAX. Pipeline inputs and outputs, as well as pipeline executables,
         are represented by logical identifiers.

         Various APIs are provided by Pegasus in order for the user to generate the
        DAX using their preferred language of choice. Available APIs include Perl, Python and Java. 
        \subsection{Pipeline execution}
        With the effecient execution of pipelines being Pegasus' main focus, there exist various components that ensure optimal processing. The Mapper's responsability is to map
        abstract workflow tasks to available resources. In order for the Mapper to successfully convert an abstract workflow into an execution workflow, it requires details on the execution
        environment. The Mapper will also query various catalogs to determine the location of executables and input and output files. These catalogs can be either file-based or in
        database format. With all the necessary information provided, the Mapper is equipped to convert abstract tasks into executable tasks. Data nodes are also added for input and output
        staging. Should the data required for a task already exist in the catalog, the task and all its parents will be removed by the Mapper from the workflow. This functionality is optional,
        and in some instances, it may be found that recomputation is preferable. If the executables required by the task are not pre-installed on the node, the Mapper will query the 
        \textit{Transformation Catalog} for the stageable binaries.

        The default, Pegasus relies on the HTCondor framework as both the workflow engine and scheduler. The executable workflow, using HTCondor DAGMan, consists of nodes representing jobs.
        These jobs each have a corresponding job submit file that details how they are meant to be executed. A job that is ready for execution is released by the DAG and placed into a Condor
        queue that is managed by the HTCondor Schedd daemon. The default scheduling is done using a local HTCondor pool, but can also be achieved through built-in interfaces with remote 
        job-management protocols, or through SSH to directly submit jobs to remote resource managers.

        Using HTCondor, Pegasus is able to achieve portability. Pegasus is compatible with local, HPC, distributed and cloud execution environments. As distributed environments may
        not employ a shared filesystem, it is recommended to use HTCondor glideins to ensure that tasks are mapped to nodes where the resources are located. Should a user running their
        pipeline locally without HTCondor installed, it is possible through \textit{submit host}. Using this method, however, executes the pipeline in serial order. Should the user wish to exploit
        the pipeline's parallelism, HTCondor is required. Pegasus pipelines are not limited to a single execution environment, and can combine multiple environments to select the best resource
        configuration to execute each job.

        In order to manage tasks and their dependencies on remote resources, three different execution engines have been developed. The first and simplest execution engine is the 
        \textit{Pegasus Cluster}, which is 
        compatible with all the execution environments. It runs on a single-thread, meaning that it can only executed a single task at a time using an ordered-list of tasks.
        The \textit{Pegasus Cluster} is particularly useful for the efficient execution of short tasks and the reduction of data transfers.

        For execution environments lacking a shared filesystem, the \textit{PegasusLite} execution engine can be used. In order for the Pegasus to execute in such environments, the Mapper first
        adds data transfer nodes to the pipeline to move the data to a data staging server associated with the execution node. Once a job is assigned to a node by the remote cluster scheduler,
        \textit{PegasusLite} is deployed.

        The final execution engine provided by Pegasus is the \textit{Pegasus MPI Cluster} (PMC). This execution engine is designed for petascale systems for which Pegasus' other strategies 
        leveraging Condor Glideins are unsuitable. It employs the Master-Worker architecture, launching a single Master and many Workers, and using MPI for communication between the 
        Master and Workers. The main limitation of this strategy is that a task must use less resources than that available on a single node. Jobs are not limited to a single execution engine,
        and thus, multiple can be used in a job to optimize performance.

        As executing a workflow on remote resources can be slow due to job queuing times and data transfers, the Mapper will group short-running task together to reduce to improve runtime. 
        When tasks located at the same distance from the DAG root have similar runtimes, level-based clustering can be used. When runtimes of same-level tasks may
        vary, it may be preferable to use job runtime-based horizontal clustering, to achieve load-balancing of clusters. Alternatively, the user may opt for label-based clustering, in which
        the user can define the task clusters themselves.

        To avoid running out of disk space during data-intensive workflows, the Mapper can add cleanup nodes that remove all unecessary data after a job has completed.

        \note{Dynamic workflows permitted}

        \subsection{Debugging and provenance}
         Job monitoring, in Pegasus, is achieved through \textit{pegasus-kickstart} (kickstart). Every job within a pegasus workflow has a kickstart wrapper that collects runtime provenance
        and performance. Kickstart relays this information to the submit host after job completion.

    \section{Geneflow}
        \subsection{Pipeline description}
        \subsection{Pipeline execution}
        \subsection{Debugging and provenance}
    \section{Toil}
        \subsection{Pipeline description}
        \subsection{Pipeline execution}
        \subsection{Debugging and provenance}
    \section{Thunder}
